Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-26 paper/ms.tex
Incorporate SAR manuscript suggestions (v0).
Co-authored-by: Saul Rappaport <saul.rappaport@icloud.com>
Co-authored-by: Saul Rappaport <sar@mit.edu>

The CQV light curves cannot be explained by pure starspots.  Although
all young M dwarfs are spotted, this tends to produce flux variations
over timescales of the rotation period, $P_{\rm rot}$,
and its half-harmonic, $0.5\,P_{\rm rot}$.  With fine-tuned
viewing geometries, starspots can produce dip durations as short
as $\approx$$0.2\,P_{\rm rot}$, and the dip amplitudes in such
geometries are small
due to limb-darkening (see \citealt{2017AJ....153..152S},
Figures 37-41).  The observed dips occur over durations as
short as $0.05\,P_{\rm rot}$ with characteristic depths of a few
percent; a ``starspot-only'' scenario can be discarded for any object
with dips that are sufficiently large in depth and short in duration
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The dip durations and
amplitudes instead require
sharp geometries with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  Since some CQVs require
such geometries, and the
CQVs probably all have a single explanation, we therefore
discard the ``spot-only'' scenario.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.3cm}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
    {\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
    stacked over one month.  Gray circles are raw 2-minute data; black
    circles bin to 300 points per cycle.  Periods in hours are in the
    bottom right of each panel.  Left-to-right, the objects are LP
    12-502 (TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and
    TIC 425933644 (Sector~28).
    {\it Bottom:} Cartoon explanations for the phenomenon.  The dust clump
    scenario (lower left) and prominence scenario (lower center) both
    invoke centrifugally-supported material at the corotation radius.
    We disfavor the screen scenario (see Section~\ref{sec:intro}).
	}
	\label{fig:f1}
\end{figure*}

@@

To assess the completeness of the resulting 120-second cadence data
that are the basis of this study, we cross-matched TIC8.2
\citep{2019AJ....158..138S,2021arXiv210804778P} against the Gaia DR2 point-source catalog
\citep{2018A&A...616A...1G}.  We opted for Gaia DR2 rather than DR3
because the base catalog for TIC8 was Gaia DR2, which facilitated a
one-to-one crossmatch using the Gaia source identifiers.  This
exercise showed that for $T$$<$16 M dwarfs, the TESS 2-minute data are
$\approx$50\% complete at $\approx$50\,pc.  At $<$20\,pc,
$\gtrsim$80\% of the $T$$<$16 M dwarfs have at least one sector of
short-cadence data; at $>$100\,pc, $\lesssim$10\% of such M dwarfs
have at least one sector of short-cadence data.  Armed with this
understanding, we then used our cross-match between Gaia DR2 and TIC8
to select our stars of interest, which we defined as stars with
120-second cadence TESS light curves satisfying these four conditions:
\begin{align}
  T &< 16 \quad&(\mathrm{Bright\ for\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only})\\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})\\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}).
\end{align}
Here, $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ is the Gaia
$G$-band absolute magnitude, $\varpi_{\rm as}$ is the parallax in
units of arcseconds, and $d$ is a geometric distance defined by
inverting the parallax and ignoring any zero-point correction.  The
target sample therefore includes \nstarssearched\ M dwarfs and late-K
dwarfs, down to $T$$<$16 and out to $d$$<$150\,pc.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CQVs}.  The complete figure
			set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable}
			is available in the online journal. {\bf For internal collaboration review:
				\url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

Our search confirms that the CQV phenomenon persists for at least
$\approx$150\,Myr.  Table~\ref{tab:thetable} includes three
$\approx$150\,Myr CQVs in AB~Dor \citep{2015MNRAS.454..593B}, a
$\approx$112\,Myr old Pleiades CQV \citep{2015ApJ...813..108D}, and a
similarly-aged Psc-Eri member \citep{2020A&A...639A..64R}.  To our
knowledge, TIC~332517282 in AB~Dor was the previous record-holder for
the oldest-known CQV \citep{2019ApJ...876..127Z,2022AJ....163..144G};
at least one unambiguous CQV (EPIC~211070495) and a few other
candidates were also previously known in the Pleiades
\citep{2016AJ....152..114R}.

@@

{\it Radial velocity scatter}---We examined diagrams of the Gaia DR3
``radial velocity error'' as a function of stellar color for all
\ncqvsnodebunked\ CQVs and candidate CQVs.  Since this quantity
represents the standard deviation of the non-published Gaia RV time
series, outliers can suggest single-lined spectroscopic binarity.
These plots showed two clusters of stars, at $\lesssim$10\,\kms and
20-25\,\kms.  We there adopted a threshold of $20$\,km\,s$^{-1}$ to
flag possible single-lined spectroscopic binaries, which selected
\nrvscatterflag\ stars: TIC~405910546, TIC~224283342, and
TIC~280945693.

@@

{\it Gaia DR3 non-single stars}---Gaia DR3 included a {\tt
non\_single\_star} column that flagged known or suspected eclipsing,
astrometric, and spectroscopic binaries.  None of the stars in our CQV
sample were identified as possible binaries in this column.

{\it Multiple periodic TESS signals}---During our visual analysis of
the TESS light curves and PDM periodograms, we flagged sources with
beating light curves, and with PDM periodograms that showed multiple
periods.  We also attempted to disentangle the two signals by
subtracting the mean CQV signal over each sector, and then repeating
the phase-dispersion minimization analysis.  The results of this effort
are summarized in the secondary periods, $P_{\rm sec}$, listed in
Table~\ref{tab:thetable}.  This process yielded
\ngoodmultperiodflag/\ngoods\ high-quality CQVs with secondary
periods; \nmaybemultperiodflag/\nmaybes\ of the ambiguous sample met
the same criterion.  Of the \ngoodhighruwe\ good CQVs with RUWE$_{\rm
DR3}$ $>$ 2, \ngoodruweandmultperiod\ also showed secondary periods in
the TESS light curves.  Considering the weaker threshold of RUWE$_{\rm
DR3}$ $>$ 1.4, \ngoodweakruweandmultperiod/\ngoodweakruwe\ such CQVs
showed secondary TESS periods.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      chunk represents one TESS orbit.  Data were acquired in Sectors
      18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The
      light curve is binned to 15-minute intervals so that there are
      96 points per day, and each point is connected by a line.  Data
      gaps are not interpolated; if data are missing, nothing
      is plotted.  The red vertical lines highlight apparently
      instantaneous state changes in the shape of the dip pattern.  
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

A few objects in Figure~\ref{fig:evoln}
show clear signs of
the CQV phenomenon in one sector, and marginal or non-existent signs
in the other.  While there is subjectivity in this assessment, to our
eyes cases for which at least one sector would be flagged as
``ambiguous'' include
TIC~368129164 (Sector 23 might be labeled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labeled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
already gave this source such a label.  However,
based on the the shape evolution between Sectors 13 and 39, it is a
CQV.  Based on the fraction of sources overall that ``turned off'',
the observed shape evolution implies that CQVs have a duty cycle of
$\approx$75\%.  This type of correction is worth including in
population-level estimates of how intrinsically common CQVs are in the
low-mass stellar population \citep[e.g.][]{2022AJ....163..144G}.

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f6.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Evolution of LP 12-502} ($P$=18.6\,h) at fixed period and
			epoch over three years.  Each panel shows one (stacked) TESS
			orbit; small text denotes relative cycle number.  There are 200
			binned black points per cycle.  The TESS pointing law dictates
			the large time gaps between Cycles 64-248, 315-1233, and
			1264-1410; larger gaps tend to yield larger shape changes.  The
			dips usually evolve over tens to hundreds of cycles.  However
			cycles 1233-1264 show a dip that ``switched'' from a depth and
			duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
			cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7d.pdf}
			\includegraphics[width=0.46\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
		{\bf River plots of the LP 12-502 light curve}, showing (clockwise
		from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
		two-harmonic sinusoid has been subtracted to highlight the sharp
		dips.  In Sectors 25-26 (cycles 248-315), periods are visible at
		the fundamental period of 18.5611\,hr, as well as at faster and
		slower relative periods.  For example, at $\phi$$\approx$0.05 and
		$\phi$$\approx$0.3, the various blue ``dips'' have distinct slopes, rather than being
		exactly vertical.  Multiple simultaneous
		periods are also visible in Sectors 53 and 58-59.  White chunks
		denote missing data.
	}
	\label{fig:lpriver0}
\end{figure*}

@@

Figure~\ref{fig:lplc} shows all available data for LP 12-502 from TESS
Sectors 18, 19, 25, 26, 53, 58, and 59, split into successive orbits.
The star was observed at 120-second cadence whenever it was observable
by TESS.  We binned the light curve to 15-minute intervals for visual
clarity, and required at least one (120-second cadence) flux
measurement per bin.  Points more than 2.5$\sigma$ above the median
are drawn in gray, also for visual clarity.  Missing data are not
drawn.  Figure~\ref{fig:lp} then shows the same data, but stacked into
successive TESS orbits spanning half a lunar month each.  Finally,
Figure~\ref{fig:lpriver0} shows ``river plots'' of the same data,
split into similar intervals: the Sector 18-19 data, 25-26 data, 53
data, and 58-59 data.  The river plots are subject to one additional
processing step: we fitted and subtracted a maximum-likelihood
two-harmonic sinusoid independently from the Sector 18-19 data, 25-26
data, and 53, 58, and 59 data in order to accentuate changes in the
dip timing and structure.

@@

Based on their locations locations in the color--absolute magnitude
diagram and membership in LCC, both of the K dwarfs do indeed appear
to be high mass.  However in detail, both are subject to ambiguities
in interpretation.  The TIC~405910546 light curve has a unique shape,
suggestive of an eclipsing binary.  Independently, TIC~405910546 was
one of only \nrvscatterflag\ CQVs flagged with a Gaia DR3 radial
velocity scatter exceeding 20\,\kms.  Combined, these factors suggest
that TIC~405910546 might be a pre-main-sequence eclipsing binary; it
should be studied further to clarify this classification.  For
TIC~405754448, we believe that the source {\it is} an unresolved
binary, because RUWE$_{\rm DR3}$=6.8, and because the raw light curves
in Sectors 11, 37, and 38 show an additional $\approx$6.5\,day,
$\approx$0.3\% amplitude sinusoidal signal suggestive of an unresolved
photometric companion.  If TIC~405754448 is a K+M binary, then the
flux ratio between the primary and secondary would be expected to be
$\approx$10:1.  It would be challenging for the M dwarf to produce the
observed 12.9\,hr signal with such a flux deficit, but not impossible.
While both of these objects suggest that the CQV phenomenon may extend
up in mass to pre-main-sequence K dwarfs, more data are needed to
fully substantiate this claim.

@@

We originally classified TIC~435903839, with RUWE$_{\rm DR3}$=17.7, as
an ``ambiguous'' CQV with a 10.8\,hr period, because this period
minimized the dispersion in the phase-folded light curve.  More
careful inspection however showed that this source is an impostor:
this source is a photometric blend of two ordinary rotating stars with
$P_0$=3.60\,hr, and $P_1$=5.41\,hr, with a beat period $(P_0^{-1} -
P_1^{-1})^{-1}$ of 10.8\,hr.  This is a novel false positive scenario
for CQVs: two rapidly-rotating stars with near-integer ratios of
rotation periods.  The beat between the two rotation signals produces
the apparent CQV signal.  Such false positives can be excluded through
careful accounting of all peaks in a periodogram.  For instance,
TIC~435903839 shows a peak at 16.27\,hr, which is not an integer
multiple of the dispersion-minimizing 10.82\,hr period.

@@

Examining Figure~\ref{fig:cqvs}, it is not obvious whether CQVs as a
whole show any preference for sharper ingresses, or sharper egresses.
In some cases (e.g. TIC~425933644), the continuum itself is not
well-defined, and so the meaning of ``ingress'' and ``egress'' are not
clear.  In others, such as Sector~36 of TIC~89463560, there is a
single clear sharp ingress with an exponential egress, which could be
fitted using e.g.~a model analogous to those used for exocomets
\citep[e.g.][]{2019A&A...625L..13Z}.  The main quantities of interest
in such models would be the exponential decay time- and therefore
length-scale, as well as the impact parameter and the inferred transit
depth (and its implications for the equivalent ``radius'' of the
transiting cloud).  We briefly explored such models, until it
became clear that careful modeling of sources such as LP
12-502 merits its own in-depth study.  Connections could likely also
be made to the toroidal geometries that can be produced by outflowing
atmospheres of transiting planets
\citep[e.g.][]{2019ApJ...873...89M,2022ApJ...926..226M}.

@@

We believe that the most likely scenarios to explain CQVs are either
the dust clump scenario or the prominence scenario
(Figure~\ref{fig:f1}).  Both invoke clumpy material that would be
centrifugally supported at the corotation radius.  The prominence idea
has a longer history, based on analogy with quiescent
prominences/filaments observed to exist in the solar corona for up to
a few weeks \citep[see][]{2015ASSL..415.....V}.  In an extrasolar
context, spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
similar structures: cold, minimally ionized hydrogen clouds or
filaments that scatter chromospheric emission from the star to produce
the observed spectroscopic line variations
\citep[see][]{1989MNRAS.238..657C}.  The short-term mechanical
stability of such gas configurations is theoretically plausible
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and the
interpretation of this class of observations seems at least somewhat
secure.

>>>>

The CQV light curves cannot be explained by pure starspots.  Although
all young M dwarfs are spotted, this tends to produce flux variations
over timescales of the rotation period, $P_{\rm rot}$,
and its half-harmonic, $0.5\,P_{\rm rot}$.  With fine-tuned
viewing geometries, starspots can produce dip durations as short
as $\approx$$0.2\,P_{\rm rot}$, and the dip amplitudes in such
geometries are small
due to limb-darkening (see \citealt{2017AJ....153..152S},
Figures 37-41).  The observed dips occur over durations as
short as $0.05\,P_{\rm rot}$ with characteristic depths of a few
percent; a ``starspot-only'' scenario can be discarded for any object
with dips that are sufficiently large in depth and short in duration
\citep{2017AJ....153..152S,2019ApJ...876..127Z,2021MNRAS.500.1366K}.   The dip durations and
amplitudes instead require
sharp geometries with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  Since some CQVs require
such geometries, and the
CQVs probably all have a single explanation, we therefore
discard the ``spot-only'' scenario.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.3cm}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
    {\it Top:} Phase-folded TESS light curves of three CQVs.  Each is
    phase-folded over one month.  Gray circles are raw 2-minute data; black
    circles bin to 300 points per cycle.  Periods in hours are in the
    bottom right of each panel.  Left-to-right, the objects are LP
    12-502 (TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and
    TIC 425933644 (Sector~28).
    {\it Bottom:} Cartoon explanations for the phenomenon.  The dust clump
    scenario (lower left) and prominence scenario (lower center) both
    invoke centrifugally-supported material at the corotation radius.
    We disfavor the screen scenario (see Section~\ref{sec:intro}).
	}
	\label{fig:f1}
\end{figure*}

@@

To assess the completeness of the resulting 120-second cadence data
that are the basis of this study, we cross-matched TIC8.2
\citep{2019AJ....158..138S,2021arXiv210804778P} against the Gaia DR2 point-source catalog
\citep{2018A&A...616A...1G}.  We opted for Gaia DR2 rather than DR3
because the base catalog for TIC8 was Gaia DR2, which facilitated a
one-to-one crossmatch using the Gaia source identifiers.  This
exercise showed that for $T$$<$16 M dwarfs, the TESS 2-minute data are
$\approx$50\% complete at $\approx$50\,pc.  At $<$20\,pc,
$\gtrsim$80\% of the $T$$<$16 M dwarfs have at least one sector of
short-cadence data; at $>$100\,pc, $\lesssim$10\% of such M dwarfs
have at least one sector of short-cadence data.  Armed with this
understanding, we then used our cross-match between Gaia DR2 and TIC8
to select our stars of interest, which we defined as stars with
120-second cadence TESS light curves satisfying these four conditions:
\begin{align}
  T &< 16 \quad&(\mathrm{Amenable\ with\ TESS})\\
  \bprp &> 1.5 \quad&(\mathrm{Red\ stars\ only})\\
  M_{\rm G} &> 4 \quad&(\mathrm{Dwarf\ stars\ only})\\
  d &< 150\,{\rm pc} \quad&(\mathrm{Close\ stars\ only}).
\end{align}
Here, $M_{\rm G} = G + 5\log(\varpi_{\rm as}) + 5$ is the Gaia
$G$-band absolute magnitude, $\varpi_{\rm as}$ is the parallax in
units of arcseconds, and $d$ is a geometric distance defined by
inverting the parallax and ignoring any zero-point correction.  The
target sample therefore includes \nstarssearched\ M dwarfs and late-K
dwarfs, down to $T$$<$16 and out to $d$$<$150\,pc.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CQVs}.  The complete figure
			set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable}
			is available in the online journal. {\bf For internal collaboration review:
				\url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc from \citet{2021A&A...649A...6G}.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

@@

Our search confirms that the CQV phenomenon persists for at least
$\approx$150\,Myr.  Table~\ref{tab:thetable} includes three
$\approx$150\,Myr CQVs in AB~Dor \citep{2015MNRAS.454..593B}, a
$\approx$112\,Myr old Pleiades CQV \citep{2015ApJ...813..108D}, and a
similarly-aged Psc-Eri member \citep{2020A&A...639A..64R}.  To our
knowledge, TIC~332517282 in AB~Dor ($t$$=$$149^{51}_{-19}$\,Myr;
\citealt{2015MNRAS.454..593B}) was the previous record-holder for the
oldest-known CQV \citealt{2019ApJ...876..127Z,2022AJ....163..144G}; at
least one unambiguous CQV (EPIC~211070495) and a few other candidates
were also previously known in the Pleiades
\citep{2016AJ....152..114R}.

@@

{\it Radial velocity scatter}---We examined diagrams of the Gaia DR3
``radial velocity error'' as a function of stellar color for all
\ncqvsnodebunked\ CQVs and candidate CQVs.  Since this quantity
represents the standard deviation of the non-published Gaia RV time
series, outliers can suggest single-lined spectroscopic binarity.
These plots showed two clusters of stars, at $\lesssim$10\,\kms and
20-25\,\kms.  We therefore adopted a threshold of $20$\,km\,s$^{-1}$ to
flag possible single-lined spectroscopic binaries, which selected
\nrvscatterflag\ stars: TIC~405910546, TIC~224283342, and
TIC~280945693.

@@

{\it Gaia DR3 non-single stars}---Gaia DR3 included a {\tt
non\_single\_star} column that flagged eclipsing, astrometric, and
spectroscopic binaries.  None of the stars in our CQV sample were
identified as possible binaries in this column.

{\it Multiple periodic TESS signals}---During our visual analysis of
the TESS light curves and PDM periodograms, we flagged sources with
beating light curves, and with PDM periodograms that showed multiple
periods.  We also attempted to disentangle the two signals by
subtracting the mean CQV signal over each sector, and then repeating
the phase-dispersion minimization analysis.  The results of this effort
are summarized in the secondary periods, $P_{\rm sec}$, listed in
Table~\ref{tab:thetable}, which we required to be at least 5\%
different from the primary period.  This process yielded
\ngoodmultperiodflag/\ngoods\ high-quality CQVs with secondary
periods; \nmaybemultperiodflag/\nmaybes\ of the ambiguous sample met
the same criterion.  Of the \ngoodhighruwe\ good CQVs with RUWE$_{\rm
DR3}$ $>$ 2, \ngoodruweandmultperiod\ also showed secondary periods in
the TESS light curves.  Considering the weaker threshold of RUWE$_{\rm
DR3}$ $>$ 1.4, \ngoodweakruweandmultperiod/\ngoodweakruwe\ such CQVs
showed secondary TESS periods.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.3cm}
		\caption{
      {\bf LP 12-502 (TIC~402980664) light curve}, where each time
      segment represents one TESS orbit.  Data were acquired in Sectors
      18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The
      light curve is binned to 15-minute intervals so that there are
      96 points per day, and each point is connected by a line.  Data
      gaps are not interpolated; if data are missing, nothing
      is plotted.  The red vertical lines highlight apparently
      instantaneous state changes in the shape of the dip pattern.  
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

A few objects in Figure~\ref{fig:evoln}
show clear signs of
the CQV phenomenon in one sector, and marginal or non-existent signs
in the other.  While there is subjectivity in this assessment, to our
eyes, cases for which at least one sector would be flagged as
``ambiguous'' include
TIC~368129164 (Sector 23 might be labeled an EB),
TIC~177309964 (Sector 38 would be simply a rotating star),
TIC~404144841 (Sector 38 looks like a rotating star),
TIC~201898222 (Sector 3 looks like a rotating star),
TIC~144486786 (Sector 32 might be an RS CVn),
and
TIC~38820496 (Sector 28 might be an RS CVn).
TIC~193831684, assessed on a single-sector basis, would probably be
labeled an eclipsing binary---in fact, \citet{2021ApJ...912..123J}
already gave this source such a label.  However,
based on the shape evolution between Sectors 13 and 39, it is a
CQV.  Based on the fraction of sources overall that ``turned off'',
the observed shape evolution implies that CQVs have an on-off duty cycle of
$\approx$75\%.  This type of correction is worth including in
population-level estimates of how intrinsically common CQVs are in the
low-mass stellar population \citep[e.g.][]{2022AJ....163..144G}.

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f6.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Evolution of LP 12-502} ($P$=18.5611\,h) at fixed period and
			epoch over three years.  Each panel shows one phase-folded TESS
			orbit; small text denotes relative cycle number.  There are 200
			binned black points per cycle.  The TESS pointing law dictates
			the large time gaps between Cycles 64-248, 315-1233, and
			1264-1410; larger gaps tend to yield larger shape changes.  The
			dips usually evolve over tens to hundreds of cycles.  However
			cycles 1233-1264 show a dip that ``switched'' from a depth and
			duration of 3\% and 3\,hr to 0.3\% and 1\,hr over less than one
			cycle (cf.~Figure~\ref{fig:lplc}).
		}
		\label{fig:lp}
	\end{center}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.46\textwidth]{f7d.pdf}
			\includegraphics[width=0.46\textwidth]{f7c.pdf}
		}
	\end{center}
	\vspace{-0.4cm}
	\caption{
    {\bf River plots of the LP 12-502 light curve}, showing (clockwise
    from top-left) Sectors 18-19, 25-26, 53, and 58-59.  A
    two-harmonic sinusoid has been subtracted to highlight the sharp
    dips.  In Sectors 25-26 (cycles 248-315), periods are visible at
    the fundamental period of 18.5611\,hr, as well as at faster
    ($\phi$$\approx$0-0.07) and slower ($\phi$$\approx$0.25-0.27)
    relative periods based on the presence of blue dips with distinct
    slopes.  Multiple simultaneous periods are also visible in Sectors
    53 and 58-59.  White chunks denote missing data.
	}
	\label{fig:lpriver0}
\end{figure*}

@@

Figure~\ref{fig:lplc} shows all available data for LP 12-502 from TESS
Sectors 18, 19, 25, 26, 53, 58, and 59, split into successive orbits.
The star was observed at 120-second cadence whenever it was observable
by TESS.  We binned the light curve to 15-minute intervals for visual
clarity, and required at least one (120-second cadence) flux
measurement per bin.  Points more than 2.5$\sigma$ above the median
are drawn in gray, also for visual clarity.  Missing data are not
drawn.  Figure~\ref{fig:lp} then shows the same data, but folded for
each successive TESS orbit.  Finally, Figure~\ref{fig:lpriver0} shows
``river plots'' of the same data, split into similar intervals: the
Sector 18-19 data, 25-26 data, 53 data, and 58-59 data.  The river
plots are subject to one additional processing step: we fitted and
subtracted a maximum-likelihood two-harmonic sinusoid independently
from the Sector 18-19 data, 25-26 data, and 53, 58, and 59 data in
order to accentuate changes in the dip timing and structure.

@@

Based on their locations in the color--absolute magnitude
diagram and membership in LCC, both of the K dwarfs do indeed appear
to be high mass.  However in detail, both are subject to ambiguities
in interpretation.  The TIC~405910546 light curve has a unique shape,
suggestive of an eclipsing binary.  Independently, TIC~405910546 was
one of only \nrvscatterflag\ CQVs flagged with a Gaia DR3 radial
velocity scatter exceeding 20\,\kms.  Combined, these factors suggest
that TIC~405910546 might be a pre-main-sequence eclipsing binary; it
should be studied further to clarify this classification.  For
TIC~405754448, we believe that the source {\it is} an unresolved
binary, because RUWE$_{\rm DR3}$=6.8, and because the raw light curves
in Sectors 11, 37, and 38 show an additional $\approx$6.5\,day,
$\approx$0.3\% amplitude sinusoidal signal suggestive of an unresolved
photometric companion.  If TIC~405754448 is a K+M binary, then the
flux ratio between the primary and secondary would be expected to be
$\approx$10:1.  It would be challenging for the M dwarf to produce the
observed 12.9\,hr signal with such a flux deficit, but not impossible.
While both of these objects suggest that the CQV phenomenon may extend
up in mass to pre-main-sequence K dwarfs, more data are needed to
fully substantiate this claim.

@@

We originally classified TIC~435903839, with RUWE$_{\rm DR3}$=17.7, as
an ``ambiguous'' CQV with a 10.8\,hr period, because this period
minimized the dispersion in the phase-folded light curve.  More
careful inspection however showed that it is an impostor:
this source is a photometric blend of two ordinary rotating stars with
$P_0$=3.60\,hr, and $P_1$=5.41\,hr, with a beat period $(P_0^{-1} -
P_1^{-1})^{-1}$ of 10.8\,hr.  This is a novel false positive scenario
for CQVs: two rapidly-rotating stars with near-integer ratios of
rotation periods.  The beat between the two rotation signals produces
the apparent CQV signal.  Such false positives can be excluded through
careful accounting of all peaks in a periodogram.  For instance,
TIC~435903839 shows a peak at 16.27\,hr, which is not an integer
multiple of the dispersion-minimizing 10.82\,hr period.

@@

Examining Figure~\ref{fig:cqvs}, it is not obvious whether CQVs as a
whole show any preference for sharper ingresses, or sharper egresses.
In some cases (e.g. TIC~425933644), the continuum itself is not
well-defined, and so the meaning of ``ingress'' and ``egress'' are not
clear.  In others, such as Sector~36 of TIC~89463560, there is a
single clear sharp ingress with an exponential egress, which could be
fitted using e.g.~a model analogous to those used for exocomets
\citep[e.g.][]{2018MNRAS.474.1453R,2019A&A...625L..13Z}.  The main quantities of interest
in such models would be the exponential decay time- and therefore
length-scale, as well as the impact parameter and the inferred transit
depth (and its implications for the equivalent ``radius'' of the
transiting cloud).  We briefly explored such models, until it
became clear that careful modeling of sources such as LP
12-502 merits its own in-depth study.  Connections could likely also
be made to the toroidal geometries that can be produced by outflowing
atmospheres of transiting planets
\citep[e.g.][]{2019ApJ...873...89M,2022ApJ...926..226M}.

@@

We believe that the most likely scenarios to explain CQVs are either
the dust clump scenario or the prominence scenario
(Figure~\ref{fig:f1}).  Both invoke clumpy material that would be
magnetically confined at the corotation radius.  The prominence idea
has a longer history, based on analogy with quiescent
prominences/filaments observed to exist in the solar corona for up to
a few weeks \citep[see][]{2015ASSL..415.....V}.  In an extrasolar
context, spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
similar structures: cold, minimally ionized hydrogen clouds or
filaments that scatter chromospheric emission from the star to produce
the observed spectroscopic line variations
\citep[see][]{1989MNRAS.238..657C}.  The short-term mechanical
stability of such gas configurations is theoretically plausible
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and the
interpretation of this class of observations seems at least somewhat
secure.
